---
title: "Using the cefasMOS package"
author: "Tom Hull"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    highlight: espresso
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# How to install and use the `cefasMOS` package

## Installing

The `cefasMOS` package is in development and has not been submitted to CRAN.
You can install it directly from bitbucket (A github analog) using the `devtools` package.

The package does not handle all of it's dependencies at the moment so it is best to install the core ones first:
Other rarely used functions may require additional packages, install these later if you have problems.


```{r install, eval=F}
install.packages("devtools")


install.packages(c("ggplot2", "lubridate", "RODBC", "dygraphs", "data.table", "insol"))

devtools::install_bitbucket("betascoo8/cefasmos")
```
```{r load, results='asis', warning=F}
library("cefasMOS") # load the package
sessionInfo("cefasMOS")$otherPkgs
```

## Extracting SmartBuoy data

The `smartbuoy.timeseries` will draw a single parameter with a zoomable timeseries plot.
It will also by default fetch telemetry data when available.
See `?smartbuoy.timeseries` for more information.

```{r timeseries, cache = T, warning=F}
x = smartbuoy.timeseries("TH1", "TEMP", yr = 2012:2014)
head(x$data)
  # x$dygraph
```

To extract QC'd data or raw data use the `smartbuoy.fetch` function.
By default only data where min QC level has been reached is returned, this can be overridden.

```{r extract, cache = T}
lbdat = smartbuoy.fetch(deployment_group = "LIVBAY", after = "2014-01-01", parameters = c("FLUORS", "TOXN"))

  # save the data
write.csv(lbdat, file = "lbdat.csv", row.names = F) # row names not needed
```